---
title: "pattern matching data"
output: html_document
date: "2024-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, here, lubridate)
```

Identifying patterns that I 'm going to try to match. 

Now have population estimates and YOY survey data

```{r}
pop.est<-read_csv(here::here("pattern_matching_data","popest2023.csv"))%>%
  dplyr::select(WBIC:Year)%>%
  mutate(WBIC=as.factor(WBIC))
yoy<-read_csv(here::here("pattern_matching_data","YOY.csv"))%>%
  filter(STATE=="WI")%>%
  rename("WBIC"=MWBC)%>%
  mutate(WBIC=as.factor(WBIC))
```

```{r}
ggplot(pop.est)+
  geom_point(aes(x=Year, y=pe, color=WBIC))+
  theme(legend.position="none")

pe.summary<-pop.est%>%
  filter(Model%in%c(0,1,2,3,4))%>%
  group_by(WBIC, Year)%>%
  summarize(n=n(),
            n.mod=length(unique(Model)))
```

```{r}
wae.pe<-pop.est%>%
  filter(Model%in%c(0,1,2,3,4))

ggplot(wae.pe)+
  geom_line(aes(x=Year, y=pe, color=WBIC))+
  theme(legend.position="none")
```

```{r}
yoy.sum<-yoy%>%
  mutate(nAge0=as.numeric(AGE0WAE),
         cpue.age0=as.numeric(AGE0CPE))%>%
  group_by(WBIC, YEAR)%>%
  summarize(mean.nAge0=mean(nAge0),
            sd.nAge0=sd(nAge0),
            n=n(),
            mean.cpue=mean(cpue.age0),
            sd.cpue=sd(cpue.age0))
```
'successful' recruitment year as cpe of 6.2 age 0 walleye per km shocked. 
in miles, t his is 9.98 f ish/mile (citing Hansen et al 2015a CJFAS,  2018; Gostiaux et al 2022)

practice seems to be filtering out lakes with less than 70% coverage of shocking (for lakes under 25.75 km/16 miles) and for larger lakes, at least 16.1 km/10 miles sampled
temps need to be between 10 and 21 degrees C (50 and 70 F), and low clarity or low conductivity is excluded

```{r}
yoy.f<-yoy%>%
  mutate(smallLake=ifelse(SHORMILE<16.0003, 1, 0),
         percentShocked=MILESURV/SHORMILE,
         drop=ifelse(smallLake==1 & percentShocked<0.7, 1,
                     ifelse(smallLake==0 & MILESURV<10, 1, 0)))%>%
  filter(TEMP>50 & TEMP<70)%>%
  filter(drop==0)%>%
  mutate(age0.cpe=as.numeric(AGE0CPE))%>%
  rename("Year"=YEAR)
  
ggplot(yoy.f)+
  geom_point(aes(x=YEAR, y=age0.cpe, color=WBIC))+
  geom_hline(yintercept=10)+
  theme(legend.position="none")
  
```

```{r}
yoy.sum<-yoy.f%>%
  mutate(rec.success=ifelse(age0.cpe>10, 1, 0))%>%
  group_by(YEAR)%>%
  summarize(prop.success=sum(rec.success, na.rm=T)/n())

ggplot(yoy.sum)+
  geom_line(aes(x=YEAR, y=prop.success))
```

```{r}
yoy.sum<-yoy.f%>%
  group_by(WBIC, Year)%>%
  mutate(nSamplingEvents=n())%>%
  summarize(age0.cpe=mean(age0.cpe),
            age0.cpe.sd=sd(age0.cpe),
            COUNTY=unique(COUNTY),
            LAKE=unique(LAKE),
            AREA=unique(AREA),
            WDNRTYPE=unique(WDNRTYPE),
            CODE=unique(CODE),
            CRNTCODE=unique(CRNTCODE))%>%
  ungroup()

yoy.wbic.year<-yoy.f%>%
  group_by(WBIC, Year)%>%
  mutate(nEvents=n())

wae.pe$WBIC_year<-paste(wae.pe$WBIC, wae.pe$Year, sep="_")
yoy.sum$WBIC_year<-paste(yoy.sum$WBIC, yoy.sum$Year,  sep="_")

yoy.error<-yoy.sum%>%
  group_by(WBIC_year)%>%
  mutate(n=n(),
         id=row_number())%>%
  # THERE ARE DUPLICATE ROWS UGH
  mutate(drop=ifelse(n==2 & id==2, 1, 0))%>%
  filter(drop==0)
  

data.join<-wae.pe%>%
  left_join(yoy.sum[,c("WBIC","Year","age0.cpe","COUNTY","LAKE","AREA","WDNRTYPE",
                     "CODE","CRNTCODE")], by=c("WBIC","Year"))

WBICs<-unique(data.join$WBIC)
  
```

411 lakes with walleye YOY survey and PE data. now see if I can get fi shing effort

```{r}
library(wdnr.fmdb)

creel_counts<-get_creel_counts(wbic=WBICs[1:100])

creel_counts2<-get_creel_counts(wbic=WBICs[101:200])

creel_counts3<-get_creel_counts(wbic=WBICs[201:300])

creel_counts4<-get_creel_counts(wbic=WBICs[301:411])

all_counts<-rbind.data.frame(creel_counts, creel_counts2, creel_counts3, creel_counts4)


creel_eff<-sum_monthly_creel_effort(all_counts)

annual_effort<-creel_eff%>%
  group_by(wbic, year)%>%
  summarize(annual.fishing.effort=sum(total.effort))%>%
  ungroup()
```
```{r}
pe.yoy.effort<-data.join%>%
  rename("year"=Year,
         "wbic"=WBIC)%>%
  mutate(wbic=as.numeric(as.character(wbic)))%>%
  left_join(annual_effort, by=c("wbic","year"))

write.csv(pe.yoy.effort, here::here("pattern_matching_data","pe.yoy.effort.csv"))
```


```{r}
complete<-pe.yoy.effort%>%
  filter(!is.na(annual.fishing.effort) & !is.na(pe) & !is.na(age0.cpe))

# 225 lakes with full 'set' of data in at least 1 year
```


So now we have walleye population estimates, age 0 CPEs (recruitment), and annual fishing effort for a bunch of lakes in WI. Looking for some general correlations now: 

```{r}
pe.yoy.effort.f<-pe.yoy.effort%>%
  filter(year>1980)

ggplot(pe.yoy.effort)+
  geom_point(aes(x=pe, y=annual.fishing.effort, color=as.factor(wbic)))+
  theme(legend.position="none")
```
```{r}
ggplot(pe.yoy.effort)+
  geom_point(aes(x=pe, y=age0.cpe, color=as.factor(wbic)))+
  facet_grid(CODE~.)+
  theme(legend.position="none")

nr<-filter(pe.yoy.effort, CODE=="NR")

ggplot(nr)+
  geom_point(aes(x=pe, y=age0.cpe, color=as.factor(wbic)))+
  theme(legend.position="none")


```
```{r}
ggplot(pe.yoy.effort)+
  geom_point(aes(x=AREA, y=annual.fishing.effort, color=as.factor(wbic)))+
  theme(legend.position="none")

```

```{r}
ggplot(pe.yoy.effort)+
  geom_point(aes(x=AREA, y=pe, color=as.factor(wbic)))+
  theme(legend.position="none")

```

